\textbf{Designing a Security Management Layer with Copy-on-Write for CDN Assets}

### 2016-02-23 tex/paper.tex
Fixed the small paper changes

%}}}
\title{\textbf{Designing a Security Management Layer for CDN Assets}}
\author{Lukas Klingsbo}

@@

\fix Clarify problem definition and research questions
\newpage
\section{Background}
This section gives the reader some background of the company which the thesis was done at and
explanation of the core concepts which are exerted.

@@

Its principle is that when processes or nodes share data in between each other, the data is not
copied until one of the processes makes changes to it. This is an optimisation as the processes does
not have to send or copy all of the related data that is in memory, rather they only have to send
pointers to the data. After many Copy-on-Write's a complex tree structure can be built up, but
optimisations can be done to simplify that structure~\cite{COPYONWRITE2}.

@@

\subsubsection{Abbreviations}
\subsubsubsection{JPF}
Java Path Finder - It was developed by NASA and in 2005 they released it under an open source
licence, which made more people contribute to the project. JPF is usually used for doing model
checking of concurrent programs to easily find for example race conditions and dead locks.

\subsubsubsection{CDN}
Content Distribution/Delivery Network - Replicates content to several servers, usually spread out
geographically. Once a request is made, the network serves content from the server closest to the
requester.

@@

\subsubsection{Terms}
\subsubsubsection{Snapshot}
A snapshot is a way to record the full state of a system at a specific time. The term comes from
photography where a photo can be seen as the state of what the photo is of, at a certain time.
Snapshots should not be confused with a full copy of a system, or part of a system, as full copies 
can be used as backups meanwhile snapshots are not very effective means of backup in the case of 
data corruption. It is not effective against data corruption as snapshots usually still refer to 
unchanged data that is still a part of the system~\cite{SNAPSHOT}.

\subsubsubsection{Eventual Consistency}
The expression eventual consistency is a term used for describing how data is processed. It is often
used in the context of database systems. Eventual consistency means that stored data is guaranteed 
to be consistent in any one moment, but at some point the data will eventually converge to being 
consistent~\cite{EVENTUAL}.

The effect of this is that persistent storages, especially distributed databases, can
operate faster as there is no need for locks and transactions for any operations. As no locks or
transactions are used, operations on the same entities can be done before earlier operations have
propagated to all nodes. This can mean that the system ends up in a state where there is a conflict
that needs to be handled, for example two operations that wants to change the same data at once. To
solve this the system needs to have defined which operation that will take precedence and this is
usually called conflict or sibling resolution.

@@

\newpage
\section{Model} \label{sec:model}
The model for this work should show how the data can not be accessed or modified by unauthorized
users and how the defined integrity of the data is always kept in the Perius system.

@@

\subsection{Approach}
As this work only presents an informal model of how the system is designed it can not be regarded as
a proof for the actual implementation of the system to be flawless. The model should how ever give a 
strong idea of the soundness of the design of the system.

The Copy-on-Write system is only used for files as all of the other objects in the system is
basically just meta data and not classed as important, from an data integrity point of view, as the 
actual files themselves.

@@

% The last one means that no users can delete actual files

@@

\subsection{Interaction of Multiple Accesses} \label{sec:multiple_access}
\fix Rename consequence?
\\
As a result of consequence~\ref{eq:contentupdate} and~\ref{eq:containerupdate} in
Section~\ref{sec:conseq} there can be race conditions where the last write wins. This could be
solved by locks or transactions, but as updates are not based on each other, the trade-off for this
inconsistency to scalability and speed is intentional. However as can also be seen in
consequence~\ref{eq:fileupdate}, if a file is removed from a content that file is not removed from
the set of files, even though its modification process is not the last one to write to the content.
This is necessary as a file can be referenced from several content and the removal process can not
atomically ensure that the file is not referred to by another content. This results in a lot of
files in the system which are not referenced from anywhere. This can be solved by either making each
file keep track of how many times it is referred to and also make it possible to check this number
and delete the file in an atomic fashion. An easier approach would be to have a garbage collector
which locks parts of the system momentarily meanwhile it removes files which are not referenced.

@@

JPF was used to test the idea and state transitioning of the model. It's a very simplified
version of the real system that still contains all the important Copy-on-Write core concepts and the 
assumptions that have been made in the model. This simplified version was then used to automatically
test for unwanted quirks. It is not a proof that the model works, but it is very exhaustive in its
testing.

@@

\par \textbf{Project} \\
A project is what is created to contain all content and containers related to a real project. Files
can be changed within a project and the system can contain several projects and their virtual
content are completely disjoint.\\

@@

Each collection in the persistent storage was emulated by using the built-in ConcurrentHashMap type.
Each client was represented by a thread and each action taken by the client was randomised. The id
hashes which MongoDB is using for each entity was imported from the mongo-java-driver-2.13.3 and
each object had its own id, generated in the same fashion as the real implementation is using,
randomly generated by the ObjectId class to minimise collisions that is.  Furthermore no locking or
transactions were used and the threads were running fully concurrently, without any sleep
statements.

@@

JPF checked each permutation of states that the threads can end up in, the result of the run can be
seen in Listing~\ref{JPFRESULT}.

@@

\subsection{Findings}
\fix Add more findings
\subsubsection{Garbage Collection}
For the system to reach the decided eventual consistency~\cite{KLINGSBO}, garbage collection will be
needed. As of the conclusion in Section~\ref{sec:multiple_access} from Section ~\ref{sec:conseq} 
there will be objects in the system that can not be reached and should therefore be cleaned up for 
them not to cause negative performance effects on the system.

@@

\newpage 
\section{Implementation} \label{sec:implementation} 
This chapter deals with the second part of this work, namely the implementation of the Perius system
based which is based on the model described in Section~\ref{sec:model}. As this work is not intended
as an instructions manual, more information about how to deploy and use Perius can be found on
\url{http://perius.se}. This chapter's focus is instead on how the decisions for the implementation
were made, how well the system would scale and the resulting security implications.

@@

In the current system you can only upload a file once as there will be a collision in the upload
otherwise, as the old and the new file will have the same MD5 hash, which is by design as duplicate
content just wastes space on the CDN. The problem with this is that the current interface does not
handle it very well as one file can not be shown or uploaded in two different places in the virtual
filesystem.

\subsubsection{Problem description}
As the current system does not offer proper security measurements, is lacking a lot of features that
is needed and does not scale very well, a new system should be developed. This work is about
examining a way of implementing Copy-on-Write in a high level system like this, which should solve
the scalability problem and make it possible to implement wanted features like snapshots, cloning
and concurrent modifications of content.

@@

\subsection{Related Technologies}
\subsubsection{React}
React is a JavaScript library for building user interfaces. React uses both its own virtual DOM and
the browser's, this makes it able to efficiently update dynamic web pages after a change of state
through comparing the old virtual DOM with the resulting virtual DOM after the state change and then
only update the browser's DOM according to the delta between the virtual DOMs~\cite{REACT}. React
can be seen as the system for handling views in front-ends implementing a MVC
(Model-View-Controller) architecture.

@@

\subsubsection{Scala} 
Scala is a multi-paradigm programming language. It most commonly runs on the JVM and compared to
Java it supports most functional programming features at the same time as it supports object
oriented programming~\cite{SCALA}.

@@

\subsubsection{MongoDB}
MongoDB is a document-oriented database which means that it does not have the concept of rows as
normal relational databases has. Instead each entity in the database is stored as a document which
is not fixed to a predefined table structure~\cite{MONGODB}. MongoDB lacks the support for joins to
improve its possibility to scale, which can be a big down side to some applications containing the
need for such logic.

\subsection{Methods for determining\\implementation details}
This chapter introduces the different methods used to determine how the new system should be
implemented, which DBMS it should use and how the estimation of long term scaling was done.

\subsection{Snapshot functionality}
\fix Structure to compare snapshot systems and conclude how Perius snapshot system was designed
\subsubsection{Copy-on-Write}
\label{sec:copy-on-write}
To efficiently create snapshots of a system Copy-on-Write can be used to make it possible to create
snapshots in O(1)\cite{BTRFS}, this is due to the fact that to create a snapshot in a system using
Copy-on-Write you only need to reference the current nodes in the tree and make sure that they are
not removed, see Figure~\ref{fig:btrfs_tree}.

As the persistent storage, used in this implementation (Section~\ref{sec:persistent_storage}), does not
implement transactions or locks a lot of different problems can occur when several clients are
working on the same data set at the same time. Such problems could be race conditions and
determining the happened-before relation. In this work this problem is solved by implementing
Copy-on-Write.
\fix Move last paragraph

@@

\subsubsection{Snapshot functionality of Perius}
In Perius snapshots and clones are not taken in the fashion which Btrfs uses, which can be seen in
Figure~\ref{fig:btrfs_tree}. As Perius does not have the tree structure pre-built and each node is
instead stored in a flat storage space, such operation would be too computationally expensive as
trees would have to be merged when collisions occur, due to the non-blocking nature of the
application. Instead this implementation makes a full copy of the meta-data of the tree, but still
refers to the same binary files until they are changed, which results in the creation of a new node.

@@

The service features a virtual file structure over the assets that has been stored, snapshots,
security management of whole containers as well as individual files, audit and auth logging, multi
project support and a modular design for persistent storage.

@@

\subsubsection{Run-time Complexity of Operations}
The runtime complexities of the operations in this section are quite intuitive, that is why they 
are only explained, and not proven to fulfil the complexities which are claimed. 
%To prove them, methods like the master theorem~\cite{ALGORITHMS} or \dots could be used.

@@

The actual storage of the data is represented by documents in a flat document storage, which is in
this case MongoDB. A very simplified version of how the data is stored can be seen in
Figure~\ref{fig:storage}. One important point that differs in the example and in the actual storage
is that GUIDs are used for identifiers and not incremented numbers, as that would not work in
distributed environments.

As the data is not stored as a tree, but rather as separate documents loosely referring to each
other, the run-time complexity is not obvious and that is what is discussed in this section.\\

\par \textbf{GET single entity} \\
When the back-end is handed a GET request it looks at what type it is and returns the document with
the correct id from the corresponding collection. As the collections are hash indexed on id this
operation can be made in constant time, O(1) and thus $\theta(1)$. \\

@@

\par \textbf{POST/Create entity} \\
Creating a new entity in the system is a constant time operation for containers and content as they
only have to be inserted as new documents in MongoDB, which is a O(1)
operation~\cite{MONGOPERFOMANCE}. Even though the insertion itself has constant run-time the index
needs to be updated or some times even rebuilt, which can slow the time before the document is
accessible in constant time. When inserting files they need to be chunked in to pieces and inserted
into GridFS, this results in n/s insertions where n is the size of the file and s is the chunk size.
The insertion of files can then be regarded as O(n) if the size of the file is regarded, but if n is
regarded as the number of files in the system the upper bound is constant as the insertion for files
is made in the same way as for other documents.\\

@@

\par \textbf{PUT/Update entity} \\
The update operation in MongoDB completely replaces a document with another in regards to their
unique identifier. This could be seen as a deletion of the old document and an insertion of the new
document, but with the same id, which makes it unnecessary to update the index for the unique id
(other indices might have to be updated though). As both deletion and insertion are constant time
operations, update is also O(1).\\

@@

This implementation is far from as efficient as the other Copy-on-Write systems described in
Section~\ref{sec:copy-on-write} in most aspects, but more efficient in some. As the implementation
is built upon MongoDB as persistent storage and not a pure tree structure, single nodes can be
fetched in O(1) but when querying for subtrees they need to be built first, which takes O(log(n)),
where n is the number of nodes in the subtree.

\subsubsection{Persistent storage} \label{sec:persistent_storage}
\subsubsubsection{MongoDB}
MongoDB was chosen as the persistent storage because of its quick lookups and because of its
internal storage format called BSON, which is very similar to JSON which the API is using. As the
formats are similar, the process of marshalling and unmarshalling becomes quite easy between the
core code, MongoDB instance and REST interface.  The second reason was that if the system needs to
scale in the future it is very easy to distribute MongoDB and if needed the system can easily be
migrated to Reactive Mongo, which is an asynchronous and non-blocking driver for MongoDB and can
therefore make the system scale even further~\cite{REACTIVEMONGO}.

All files are also stored directly in MongoDB with the help of GridFS. GridFS chunks the files 
according to the size limit of MongoDB objects, which is currently 4MB. The advantage of this is 
that backups of the Perius state is easily done through a database backup, no separate files 
needs to be backed up. Another advantage that is given by this is that you can retrieve specific 
ranges of a file, although that advantage is not needed in the Perius implementation. For
scalability this could be used to retrieve different parts of a file from different servers, 
normal load balancing would probably work better in a system like Perius where no extremely large 
files are expected to be stored.

The disadvantage of using the GridFS approach is that when using a non-distributed database 
it is slower to read and write to the database than reading or writing directly to the filesystem. 
Another disadvantage is that to access the files it is needed to go through the database layer 
in some way, instead of accessing the filesystem.

@@

REST was chosen as only basic CRUD operations needs to be performed and because the BSON format
which is used in MongoDB is almost identical ~\cite{BSON} to the standardised JSON format which is
usually used by RESTful services~\cite{JSON}.

@@

\newpage
\subsection{Load Testing} \label{sec:load_testing}
Wrk was used to load test the back-end. It is a multi-threaded benchmarking tool for HTTP which can
create large loads. The testing was done locally on a virtual server having 6 CPU cores and 48GB of
RAM. The results that will be analysed are latency and throughput of the server as that will give a 
fair judgement of how well the system can perform.

\subsubsection{GET requests} \label{sec:GET_REQUESTS}
To be able to determine what is the bottleneck of the back-end three different types of \textit{GET
requests} are performed. The first one tries to maximise the number of requests that spray-can (HTTP
server) can handle with the given specifications and therefore the server just returns 200 (OK). The
second request requests the simplest API call which involves getting all currently stored projects
from the MongoDB instance. As this involves the server answering with some payload, the third
request is simply the same amount of static payload as the second request but without involving any
connections to MongoDB.
\par
The typical requests with Wrk in this work will look similar to this:
\begin{lstlisting}[frame=single]
wrk -t100 -c100 -d10s http://perius:8000/ok
\end{lstlisting}
where -t100 means that it uses 100 threads, -c100 that it emulates 100 clients requesting over and
over and -d100s is the amount of time that the load test will run, in this case 100 seconds.

@@

Listing~\ref{OKREQUEST} shows that the HTTP server can answer about 100K requests/s when not
involving any payload other than status code 200 (OK). When comparing that to the request which
involved the server responding with a small (100 Byte) payload (Listing~\ref{STATICREQUEST} one can
see that it is about 10\% slower to send some more data, \textasciitilde90K requests/s. However when
comparing that to the requests that needed database access it is obvious that the HTTP server can
handle all the load that it needs to, as Listing~\ref{DBREQUEST} shows that to fetch all documents
in a collection (in this case only one) is dramatically slower, the throughput was only
\textasciitilde7000 requests/s.

\par These are very simple requests, to evaluate how much load the application can handle in its
current state a more complex but common request has to be analysed. The most complicated and still
common request that system is receiving is requests of a full project tree. This request is
computationally expensive as the tree is not stored directly in the database but has to be built
from the id and parent id of each container and content, a simulated project tree of similar size to
the ones stored in the current management system reaches around 25KB in uncompressed size.

@@

\par
Another simple optimisation that was added was indices on the parent ids, which are heavily queried
when building the project trees.

@@

\par
As the project tree is built by querying documents parent id an optimisation that could be done was
to create an index of the documents parent ids. This drastically improved the size of the tree that
the server was able to handle within the time out, going from being able to handle 4000 documents in 
one tree to at least 50000 documents, which the browser instead will have problems displaying, 
represented in the UI, without lag.

\par
After adding indices, the building of the project tree was able to finish 50 times faster and the
server was able to serve 2100 requests/s, see Listing~\ref{INDEXREQUEST}. If using the same
approximated average that was used in connection to Listing~\ref{TREEREQUEST} (30 requests/s), the
maximum number of clients that one node can handle will be >60000, without taking the web servers
limits or content posting into consideration.

@@

\subsubsection{Audit logs}
Audit logs are used in the system to keep track of which users that perform which actions. In the
event that somebody, consciously or unconsciously, are leaking private CDN content from within the
organisation using Perius the audit trail can be used to find which LDAP account that was
responsible for the leak.

@@

\newpage
\section{Discussion}
\subsection{Persistent Storage}
More research could have been done in the choosing of persistent storage, as mentioned in 
\textit{``NoSQL: Moving from MapReduce Batch Jobs to Event-Driven Data Collection''}
~\cite{KLINGSBO} many applications that choose NoSQL databases as their persistent storage 
actually don't need it. This is most likely true for Perius too, it would have been efficient 
enough generating the project trees from an indexed SQL database with foreign keys. On the 
other hand it was quite nice having the BSON documents for insertion as they were so similar 
the accepted JSON format from the REST service. As Perius most common operations involve 
modifying the project tree, a graph database should have been researched too.

@@

\newpage
\section{Summary}
\subsection{Conclusions}
The goal of this thesis was to see whether it was feasible to use Copy-on-Write in a high level
application (See Figure~\ref{fig:stack}). As the implementation (see
Section~\ref{sec:implementation}) made as a part of this thesis project is already replacing its
predecessor the simple conclusion is that it definitely is possible. In the beginning of the project
thoughts were on having every element being operated upon in a copy-on-write fashion but this was
later narrowed down to only have the most important part, the files, as Copy-on-Write. This was due
to that the conclusion that it did not matter if the other elements were resolved in a
last-write-wins manner when modified.

\par If this application would be distributed with several persistent storage nodes like MongoDB,
the application would not always be in a global consistent state as there would not be any global
locks.  This could theoretically cause some inconvenience for the user but in all real world tests
no users have noticed it. The theoretical inconvenience for the user was a trade-off made so that
the application could scale on the width almost endlessly, especially if the issue with building the
project trees is (as mentioned in Section~\ref{sec:load_testing}) solved.

>>>>

%}}}
\title{\textbf{Designing a Security Management Layer with Copy-on-Write for CDN Assets}}
\author{Lukas Klingsbo}

@@

\fix Clarify problem definition and research questions
\newpage
\section{Background}
This section gives the reader some background of the company where the thesis was written and
implementation was made. It also gives some brief explanation of the core concepts which are 
exerted throughout this work.

@@

Its principle is that when processes or nodes share data in between each other, the data is not
copied until one of the processes makes changes to it. This is an optimisation as the processes does
not have to send or copy all of the related data that is in memory, rather they only have to send
pointers to the data. After many Copy-on-Writes a complex tree structure can be built up, but
optimisations can be done to simplify that structure~\cite{COPYONWRITE2}.

@@

\subsubsection{Abbreviations}
\subsubsubsection{CDN}
Content Distribution/Delivery Network - Replicates content to several servers, usually spread out
geographically. Once a request is made, the network serves content from the server closest to the
requester.

@@

\subsubsubsection{JPF}
Java Path Finder - It was developed by NASA and in 2005 they released it under an open source
licence, which made more people contribute to the project. JPF is usually used for doing model
checking of concurrent programs to easily find for example race conditions and dead locks.

\subsubsection{Terms}
\subsubsubsection{Eventual Consistency}
The expression eventual consistency is a term used for describing how data is processed. It is often
used in the context of database systems. Eventual consistency means that stored data is guaranteed 
to be consistent in any one moment, but at some point the data will eventually converge to being 
consistent~\cite{EVENTUAL}.

The effect of this is that persistent storages, especially distributed databases, can
operate faster as there is no need for locks and transactions for any operations. As no locks or
transactions are used, operations on the same entities can be done before earlier operations have
propagated to all nodes. This can mean that the system ends up in a state where there is a conflict
that needs to be handled, for example two operations that wants to change the same data at once. To
solve this the system needs to have defined which operation that will take precedence, or how the
concurrent operations are mixed, and this is usually called conflict or sibling resolution.

@@

\subsubsubsection{Snapshot}
A snapshot is a way to record the full state of a system at a specific time. The term comes from
photography where a photo can be seen as the state of what the photo is of, at a certain time.
Snapshots should not be confused with a full copy of a system, or part of a system, as full copies 
can be used as backups meanwhile snapshots are not very effective means of backup in the case of 
data corruption. It is not effective against data corruption as snapshots usually still refer to 
unchanged data that is still a part of the system~\cite{SNAPSHOT}.

\newpage
\section{Model} \label{sec:model}
The model for this work should show how the data can only be accessed or modified by authorized
users and how the defined integrity of the data is always kept in the Perius system.

@@

\subsection{Approach}
As this work only presents an informal model of how the system is designed it can not be regarded as
a proof for the actual implementation of the system to be flawless. The model should however give a 
strong idea of the soundness of the design of the system.

The Copy-on-Write system is only used for files as all of the other objects in the system is
basically just meta data and not classed as important, from a data integrity point of view, as the 
actual files themselves.

@@

\fix The last one means that no users can delete actual files, probably a bit confusingly written, 
or even wrong?

@@

\subsection{Interaction of Multiple Accesses} \label{sec:multiple_access}
\fix Rename consequence?
\\
As a result of consequence~\ref{eq:contentupdate} and~\ref{eq:containerupdate} in
Section~\ref{sec:conseq} there can be race conditions where the last write wins. This could be
solved by locks or transactions, but as updates are not based on each other, the trade-off for this
inconsistency to scalability and speed is intentional. However, as can also be seen in
consequence~\ref{eq:fileupdate}, if a file is removed from a content that file is not removed from
the set of files, even though its modification process is the last one to write to the content.
This is necessary as a file can be referenced from several content and the removal process can not
atomically ensure that the file is not referred to by another content. This results in a lot of
files in the system which are not referenced from anywhere. This can be solved by either making each
file keep track of how many times it is referred to and also make it possible to check this number
and delete the file in an atomic fashion. An easier approach would be to have a garbage collector
which locks parts of the system momentarily meanwhile it removes files which are not referenced.

@@

JPF was used to test the idea and state transitioning of the model. It's a very simplified
version of the real system that still contains all the important Copy-on-Write concepts and the 
assumptions that have been made in the model. This simplified version was then used to automatically
test for unwanted quirks. It is not a proof that the model works, but it is very exhaustive in its
testing.

@@

\par \textbf{Project} \\
A project is what is created to contain all content and containers related to a real project. Files
and meta-data can only be changed within a project and the system can contain several projects 
that have their virtual content and containers completely disjoint.\\

@@

Each collection in the persistent storage was emulated by using the built-in ConcurrentHashMap type.
Each client was represented by a thread and each action taken by the client was randomised. The id
hashes which MongoDB is using for each entity was imported from the mongo-java-driver-2.13.3 and
each object had its own id, generated in the same fashion as the real implementation is using,
randomly generated by the ObjectId class to minimise collisions that is.  Furthermore no locking or
transactions were used and the threads were running fully concurrently and without any sleep
statements.

@@

JPF checked each permutation of states, to a certain depth, that the threads can end up in and the 
result of the run can be seen in Listing~\ref{JPFRESULT}.

@@

\subsection{Findings}
\fix Add more findings
\subsubsection{Garbage Collection}
For the system to reach a clean state and the decided eventual consistency~\cite{KLINGSBO}, garbage
collection will be needed. As of the conclusion in Section~\ref{sec:multiple_access} from Section
~\ref{sec:conseq} there will be objects in the system that are unreferenced and can not be reached
and should therefore be cleaned up for them not to cause negative performance effects on the system.

@@

\newpage 
\section{Implementation} \label{sec:implementation} 
This chapter deals with the second part of this work, namely the implementation of the Perius
system. The implementation is based on the model described in Section~\ref{sec:model}. As this work
is not intended as an instructions manual, more information about how to deploy and use Perius can
be found on \url{http://perius.se}. This chapter's focus is instead on how the decisions for the
implementation were made, how well the system would scale and the resulting security implications.

@@

In the current system you can only upload a file once as there will be a collision in the upload
otherwise, as the old and the new file will have the same MD5 hash, which is by design as duplicate
content just wastes space on the CDN. The problem with this is that the current interface does not
handle it very well as one file can not be shown or uploaded in two different places in Battle
Binary's virtual filesystem.

\subsubsection{Problem description}
As the current system does not offer proper security measurements, is lacking a lot of features that
are needed and does not scale very well, a new system should be developed. This work is about
examining a way of implementing Copy-on-Write in a high level system like this, which should solve
the scalability problem and make it possible to implement wanted features like snapshots, cloning
and concurrent modifications of content.

@@

\subsection{Related Technologies}
\subsubsection{React}
React is a JavaScript library for building user interfaces. React uses both its own virtual DOM and
the browser's normal DOM, this makes it able to efficiently update dynamic web pages after a change
of state through comparing the old virtual DOM with the resulting virtual DOM after the state change
and then only update the browser's DOM according to the delta between the virtual DOMs~\cite{REACT}.
React can be seen as the system for handling views in front-ends implementing a MVC
(Model-View-Controller) architecture.

@@

\subsubsection{Scala} 
Scala is a multi-paradigm programming language. It most commonly runs on the JVM and as apart from
Java it supports most functional programming features at the same time as it supports object
oriented programming~\cite{SCALA}.

@@

\subsubsection{MongoDB}
MongoDB is a document-oriented database, which means that it does not have the concept of rows as
normal relational databases has. Instead each entity in the database is stored as a document which
is not fixed to a predefined table structure~\cite{MONGODB}. MongoDB lacks the support for joins to
improve its possibility to scale, which can be a big down side to some applications containing the
need for such logic.

\subsection{Methods for determining\\implementation details}
This chapter introduces the different methods used to determine how the new system should be
implemented, which DBMS it should use and how the estimation of long term scaling was done.
\\
\fix Fix this section, move things here and fill out with missing details

\subsection{Snapshot functionality}
\fix Structure to compare snapshot systems and conclude how Perius snapshot system was designed
\subsubsection{Copy-on-Write}
\label{sec:copy-on-write}
To efficiently create snapshots of a system, Copy-on-Write can be used to make it possible to create
snapshots in O(1)\cite{BTRFS}. This is due to the fact that to create a snapshot in a system using
Copy-on-Write you only need to reference the current nodes in the tree and make sure that they are
not removed, see Figure~\ref{fig:btrfs_tree}.

As the persistent storage, used in this implementation (Section~\ref{sec:persistent_storage}), does not
implement transactions or locks a lot of different problems can occur when several clients are
working on the same data set at the same time. Such problems could be race conditions or
determining the happened-before relation. In this work this problem is solved by implementing
Copy-on-Write.
\fix Move last paragraph

@@

\subsubsection{Snapshot functionality of Perius}
In Perius snapshots and clones are not taken in the fashion which Btrfs uses, which can be seen in
Figure~\ref{fig:btrfs_tree}. As Perius does not have the tree structure pre-built and each node is
instead stored in a flat storage space, such operation would be too computationally expensive as
trees would have to be merged when collisions occur, due to the non-blocking nature of the
application. Instead this implementation makes a full copy of the meta-data of the tree, but still
refers to the same binary files until they are changed.

@@

The service features a virtual file structure over the assets and meta-data that have been stored,
as can be seen in Figure~\ref{fig:frontend}. It also features snapshots, security management of
whole containers as well as individual files, audit and auth logging, multiple separate project 
and a design which allows for an interchangeable persistent storage.

@@

\subsubsection{Run-time Complexity of Operations} \label{sec:run_time}
The runtime complexities of the operations in this section are quite intuitive, that is why they 
are only explained, and not proven to fulfil the complexities which are claimed. 
%To prove them, methods like the master theorem~\cite{ALGORITHMS} or \dots could be used.

@@

The actual storage of the data is represented by documents in a flat document storage, which is in
this case MongoDB. A very simplified version of how the data is stored can be seen in
Figure~\ref{fig:storage}, however in reality it not quite so table-like. One important point that 
differs in the example and in the actual storage is that GUIDs are used for identifiers and not 
incremented numbers, as that would not work in distributed environments.

The data is not stored as a tree (see Example~\ref{fig:storage}, but rather as separate documents 
loosely referring to each other. How the different operations perform their actions and what 
implications these actions have on the operations run-time complexities are not obvious and that 
is what is discussed in this section.\\

\par \textbf{GET single entity} \\
When the back-end is handed a GET request it looks at what type of entity that is requested and
returns the document with the correct id from the corresponding collection. As the collections are
hash indexed on id this operation can be made in constant time, O(1) and thus $\theta(1)$. \\

@@

\par \textbf{POST/Create entity} \\
Creating a new entity in the system is a constant time operation for containers and content as they
only have to be inserted as new documents in MongoDB, which is a O(1)
operation~\cite{MONGOPERFORMANCE}. Even though the insertion itself has constant run-time the index
needs to be updated or sometimes even rebuilt, which can slow down the time before the document is
accessible in constant time. When inserting files they need to be chunked into pieces and inserted
into GridFS, this results in n/s insertions where n is the size of the file and s is the chunk size.
The insertion of files can then be regarded as O(n) if the size of the file is regarded, but if n is
regarded as the number of files in the system the upper bound is constant as the insertion for files
is made in the same way as for other documents.\\

@@

\par \textbf{PUT/Update entity} \\
The update operation in MongoDB completely replaces a document with another in regards to their
unique identifier. This could be seen as a deletion of the old document and an insertion of a new
one, but with the same id, which makes it unnecessary to update the index for the unique id. Other 
indices might have to be updated though, if for example the parent of the document is changed. 
As both deletion and insertion are constant time operations, update is also O(1).\\

@@

This implementation is far from as efficient as the other Copy-on-Write systems described in
Section~\ref{sec:copy-on-write} in most aspects, but more efficient in some. As the implementation
is built upon MongoDB as persistent storage and not a pure tree structure, single nodes can be
fetched in O(1) but when querying for subtrees they need to be built first, which arguable (See
Section~\ref{sec:run_time}) takes $\theta(log(n))$, where n is the number of nodes in the subtree.

\subsubsection{Persistent storage} \label{sec:persistent_storage}
\subsubsubsection{MongoDB}
MongoDB was chosen as the persistent storage because of its quick lookups and because of its
internal storage format called BSON, which is very similar to JSON which the API is using. As the
formats are similar, the process of marshalling and unmarshalling becomes quite easy between the
core code, MongoDB instance and REST interface.  The second reason was that if the system needs to
scale in the future it is very easy to distribute MongoDB and, if needed, the system can easily be
migrated to Reactive Mongo. Reactive Mongo is an asynchronous and non-blocking driver for MongoDB
and can therefore make the system scale even further~\cite{REACTIVEMONGO}.

All files are also stored directly in MongoDB with the help of GridFS. GridFS chunks the files 
according to the size limit of MongoDB objects, which is currently 4MB. The advantage of this is 
that backups of the Perius system is easily done through a database backup, no separate files 
needs to be backed up. Another advantage that is given by this is that you can retrieve specific 
ranges of a file, although that advantage is not needed in the Perius implementation. For
scalability this could be used to retrieve different parts of a file from different servers, 
normal load balancing would probably work better in a system like Perius where no extremely large 
files are expected to be stored.

The disadvantage of using the GridFS approach is that when using a non-distributed database 
it is slower to read and write to the database than reading or writing directly to a filesystem. 
Another disadvantage is that to access the files it is needed to go through the database layer 
in some way, instead of accessing the filesystem directly.

@@

REST was chosen as all operations can be defined as basic CRUD operations needs to be performed and
because the BSON format which is used in MongoDB is almost identical ~\cite{BSON} to the
standardised JSON format which is usually used by RESTful services~\cite{JSON}.

@@

\newpage
\subsection{Load Testing} \label{sec:load_testing}
Wrk was used to load test the back-end. It is a multi-threaded benchmarking tool for HTTP which can
create large loads. The testing was done locally on a virtual server having 6 CPU cores and 48GB of
RAM. The results that will be analysed are throughputs performed without breaking the set timeout 
of the server, as that will give a fair judgement of how well the system can perform.

\subsubsection{GET requests} \label{sec:GET_REQUESTS}
To be able to determine what the bottleneck of the back-end is, three different types of \textit{GET
requests} are performed. The first one tries to maximise the number of requests that spray-can (HTTP
server) can handle with the given specifications and therefore the server just returns 200 (OK). The
second request requests the simplest API call which involves getting all currently stored projects
from the MongoDB instance. As this involves the server answering with some payload, the third
request is simply the same amount of static payload as the second request but without involving any
connections to MongoDB.
\par
The typical requests with Wrk in this work will look similar to this:
\begin{lstlisting}[frame=single]
wrk -t100 -c100 -d10s http://perius:8000/ok
\end{lstlisting}
where -t100 means that it uses 100 threads, -c100 that it emulates 100 clients requesting over and
over and -d100s is the amount of time that the load test will run, in this case 100 seconds.

@@

Listing~\ref{OKREQUEST} shows that the HTTP server can answer about 100K requests/s when not
involving any payload other than status code 200 (OK). When comparing that to the request which
involved the server responding with a small (100 Byte) payload (Listing~\ref{STATICREQUEST}) one can
see that it is about 10\% slower to send some more data, \textasciitilde90K requests/s. However when
comparing that to the requests that needed database access it is obvious that the HTTP server can
handle all the load that it needs to, as Listing~\ref{DBREQUEST} shows that to fetch all documents
in a collection (in this case only one) is dramatically slower, the throughput was only
\textasciitilde7000 requests/s.

\par These are very simple requests, to evaluate how much load the application can handle in its
current state, a more complex but common request has to be analysed. The most complicated and still
common request that the system is receiving is requests of a full project tree. This request is
computationally expensive as the tree is not stored directly in the database but has to be built
from the id and parent id of each container and content, a simulated project tree of similar size to
the ones stored in the current management system reaches around 25KB in uncompressed size.

@@

@@

\par As a project tree is built by querying documents parent id an optimisation that could be done was
to create an index of the documents parent ids. This drastically improved the size of the tree that
the server was able to handle within the time out, going from being able to handle 4000 documents in 
one tree to at least 50000 documents, which the browser instead will have problems displaying, 
represented in the UI, without lag.

\par After adding indices, the building of the project tree was able to finish 50 times faster and
the server was able to serve 2100 requests/s, see Listing~\ref{INDEXREQUEST}. If using the same
approximated average that was used in connection to Listing~\ref{TREEREQUEST} (30 requests/s), the
maximum number of active clients that one node can handle will be >60000, without taking the web
servers limits or content posting into consideration.

@@

\subsubsection{Audit logs}
Audit logs are used in the system to keep track of which users that perform which actions. In the
event that somebody, consciously or unconsciously, are leaking private CDN content from within the
organisation using Perius. The audit trail can then be used to find which LDAP account that was
responsible for the leak.

@@

\newpage
\section{Discussion}
\subsection{Persistent Storage}
More research could have been done in the choosing of persistent storage, as mentioned in 
\textit{``NoSQL: Moving from MapReduce Batch Jobs to Event-Driven Data Collection''}
~\cite{KLINGSBO} many applications that choose NoSQL databases as their persistent storage 
actually don't need it. This is most likely true for Perius too, it would have been efficient 
enough generating the project trees from an indexed SQL database with foreign keys. On the 
other hand it was quite nice having the BSON documents for insertion as they were so similar 
to the accepted JSON format from the REST service. As Perius most common operations involve 
modifying project trees, a graph database should have been researched too.

@@

\newpage
\section{Summary}
\subsection{Conclusions}
The goal of this thesis was to see whether it was feasible to use Copy-on-Write in a high level
application (See Figure~\ref{fig:stack}). As the implementation (see
Section~\ref{sec:implementation}) made as a part of this thesis project is already replacing its
predecessor the simple conclusion is that it definitely is possible. In the beginning of the project
thoughts were on having every element being operated upon in a Copy-on-Write fashion but this was
later narrowed down to only have the most important part, the files, as Copy-on-Write. This was due
to that the conclusion that it did not matter if the other elements were resolved in a
last-write-wins manner when modified.

\par If this application would be distributed with several persistent storage nodes like MongoDB,
the application would not always be in a global consistent state as there would not be any global
locks.  This could theoretically cause some inconvenience for the user but in all real world
observations no users have noticed it. The theoretical inconvenience for the user was a trade-off
made so that the application could scale on the width almost endlessly, especially if the issue with
building the project trees is (as mentioned in Section~\ref{sec:load_testing}) solved.
